---
title: "Centrality indices"
author: "David Schoch"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{centrality indices}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
This vignette describes how to build different centrality indices on the basis of 
indirect relations as described in [this](indirect_relations.html) vignette. Note,
however, that the primary purpose of the netrankr package is **not** to provide a great
variety of indices, but to offer alternative methods for centrality assessment. 
Consult the tutorial and this vignette, before starting trial-and-error usage of indices.

## Building indices with the `netrankr` package

```{r setup, warning=FALSE,message=FALSE}
library(netrankr)
library(igraph)
library(magrittr)
```

The `netrankr` does, by design, not explicitly implement any centrality index. It 
does, however, provide a huge set of components to build indices.

The traditional indices degree, closeness, betweenness and eigenvector centrality 
can 

```{r standardcent}
g <- graph.empty(n=11,directed = FALSE)
g <- add_edges(g,c(1,11,2,4,3,5,3,11,4,8,5,9,5,11,6,7,6,8,
                   6,10,6,11,7,9,7,10,7,11,8,9,8,10,9,10))
V(g)$name <- 1:11

#adjacency
g %>% indirect_relations(g,type="identity")
#geodesic distances
D <- indirect_relations(g,type="geodesic")
#dyadic dependencies (as used in betweenness centrality)
B <- indirect_relations(g,type="dependencies")
#resistance distance (as used in information centrality)
R <- indirect_relations(g,type="resistance")
```