---
title: "Positional dominance in networks"
author: "David Schoch"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{positional dominance in networks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette describes the concept of positional dominance, the generalization 
of [neighborhood-inclusion](neighborhood_inclusion.html) for arbitrary network and attribute data. Additionally, 
some use cases with the `netrankr` package are given.

## Theoretical Background



More about the positional dominance and the positional approach to network analysis can be found in

> Brandes, Ulrik. (2016). Network Positions.
*Methodological Innovations*, **9**, 2059799116630650.
([link](http://dx.doi.org/10.1177/2059799116630650))

## Positional Dominance in the `netrankr` Package

```{r setup, warning=FALSE,message=FALSE}
library(netrankr)
library(igraph)
library(magrittr)
set.seed(1886) #for reproducibility
```

The function `positional_dominance` can be used to check 
In accordance with the analytic pipeline of centrality[REF HERE] we use the `%>%` operator from the 
`magrittr` package.

```{r pos dom}
g <- graph.empty(n=11,directed = FALSE)
g <- add_edges(g,c(1,11,2,4,3,5,3,11,4,8,5,9,5,11,6,7,6,8,
                   6,10,6,11,7,9,7,10,7,11,8,9,8,10,9,10))
V(g)$name <- 1:11

#neighborhood inclusion can be expressed with the analytic pipeline
D <- g %>% indirect_relations(relation="identity") %>% positional_dominance()

#without %>% operator:
# rel <- indirect_relations(g,relation="identity")
# D <- positional_dominance(rel)

identical(D,neighborhood_inclusion(g))
```

the `map` parameter of `positional_dominance` allows to distinguish between dominance
under *total heterogeneity* (`map=FALSE`) and *total homogeneity* (`map=TRUE`). 
Dominance under total homogeneity yields a ranking, if the relation is binary (e.g. connected or not).  

```{r homo and hetero}
D <- g %>% indirect_relations(relation="identity") %>% positional_dominance(map=TRUE)

comparable_pairs(D)
```


