---
title: "Uniquely ranked graphs"
author: "David Schoch"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Uniquely ranked graphs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette introduces *threshold graphs*, a class of graphs with a unique centrality ranking and
relevant functions from the `netrankr` package to work with this class of graphs. 

## Theoretical Background

A threshold graph is a graph, where all nodes are pairwise comparable by neighborhood inclusion. Formally,
$$
\forall u,v \in V: N(u) \subseteq N[v] \; \lor \; N(v) \subseteq N[u].
$$

According to [this](neighborhood_inclusion.html) vignette, it is thus clear that all centrality indices induce the same ranking on a threshold graph. More technical details on threshold graphs and results related to centrality can be found in

> Schoch, David & Valente, Thomas W., & Brandes, Ulrik. (2017). Correlations among centrality indices
and a class of uniquely ranked graphs. *Social Networks*, **50**, 46-54.([link](http://doi.org/10.1016/j.socnet.2017.03.010))

## Uniquely Ranked Graphs in the `netrankr` Package

```{r setup, warning=FALSE,message=FALSE}
library(netrankr)
library(igraph)
set.seed(1886) #for reproducibility
```
Threshold graphs on $n$ vertices can be constructed iteratively with a sequence of $0$'s and $1$'s.
For each $0$, an isolated vertex is inserted and for each $1$ a vertex that connects to all previously inserted one's. This iterative process is implemented in the `threshold_graph` function. The parameter `n` is used to set the desired number of vertices. The parameter `p` is the probability that a dominated vertex is inserted in each step. This parameter roughly equates to the density of the network.
```{r tg creation}
 g1 <- threshold_graph(500,0.4)
 g2 <- threshold_graph(500,0.05)
 
 c(round(graph.density(g1),2), round(graph.density(g2),2))
```

The class of threshold graphs includes various well-known graphs, for instance star shaped and complete networks. This graphs can be constructed with `p=0` and `p=1` respectively.
```{r tg example,fig.show='hold'}
star <- threshold_graph(6,0) 
complete <- threshold_graph(6,1)
plot(star,vertex.label=NA,vertex.color="black")
plot(complete,vertex.label=NA,vertex.color="black")
```

To check that all pairs are comparable by neighborhood inclusion, we can use the function `comparable_pairs`. The function computes the density of the underlying undirected graph induced by the neighborhood-inclusion relation.
```{r tg complete}
g <- threshold_graph(10,0.4)
P <- neighborhood_inclusion(g)
comparable_pairs(P)
```

## Correlation and Threshold Graphs
We construct a random threshold graph and calculate some standard measures of centrality that are included in the `igraph` package.
```{R tg cent}
g <- threshold_graph(100,0.1)

cent.df <- data.frame(
  degree=degree(g),
  betweenness=betweenness(g),
  closeness=closeness(g),
  eigenvector=round(eigen_centrality(g)$vector,8),
  subgraph=subgraph_centrality(g)
)
```

We expect, that all indices are perfectly rank correlated since all pairs of nodes are comparable by neighborhood-inclusion.
```{r cor tg}
cor.mat <- cor(round(cent.df,8),method="kendall")
cor.mat <- round(cor.mat,2)
cor.mat
```

We, however, obtain correlations that are not equal to one. This is due to the definition of Kendall's (tie corrected) $\tau$. 

The following cases can arise when comparing two scores of indices `x` and `y`.

* concordant: `x[i]>x[j] & y[i]>y[j]` or `x[i]<x[j] & y[i]<y[j]`
* discordant: `x[i]>x[j] & y[i]<y[j]` or `x[i]>x[j] & y[i]<y[j]`
* tied: `x[i]=x[j] & y[i]=y[j]`
* left/right ties: `x[i]=x[j] & y[i]!=y[j]` or `x[i]!=x[j] & y[i]=y[j]`

Kendall's $\tau$ considers left and right ties as a correlation reducing. Left and right ties are, however, not forbidden according to the neighborhood inclusion property. The only forbidden case are discordant pairs. That is, $N(u)\subseteq N[v]$ can not result in $c(u)>c(v)$ but it may result in $c(u)=c(v)$.

The function `compare_ranks` can be used to calculate the number of occurences for each case. 

```{R compare ranks}
comp <- compare_ranks(cent.df$degree,cent.df$betweenness)
unlist(comp)
```

Notice that there is a high number of right ties which influences the correlation if measured with Kendall's $\tau$. However, there do not exist any discordant pairs for any pair of indices.

```{r discordant}
dis.pairs <- matrix(0,5,5)
dis.pairs[1,] <- apply(cent.df,2,
                       function(x)compare_ranks(cent.df$degree,x)$discordant)
dis.pairs[2,] <- apply(cent.df,2,
                       function(x)compare_ranks(cent.df$betweenness,x)$discordant)
dis.pairs[3,] <- apply(cent.df,2,
                       function(x)compare_ranks(cent.df$closeness,x)$discordant)
dis.pairs[4,] <- apply(cent.df,2,
                       function(x)compare_ranks(cent.df$eigenvector,x)$discordant)
dis.pairs[5,] <- apply(cent.df,2,
                       function(x)compare_ranks(cent.df$subgraph,x)$discordant)
dis.pairs
```

Although Kendall's $\tau$ suggests that the correlations among indices can be low,
we see that there do not exist any discordant pairs on threshold graphs.