---
title: "Use Case: Florentine Families"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{07 use case}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette describes a typical use case of the `netrankr` packages. It contrasts
the usual centrality analysis based on indices with the dominance based assessment.
It is advisable to read the other vignettes before going through this example.
________________________________________________________________________________


## Setup
[Padget & Ansell, 1993](http://www.journals.uchicago.edu/doi/abs/10.1086/230190)

```{r setup, warning=FALSE,message=FALSE}
library(igraph)
library(netrankr)
library(magrittr)
library(ggplot2)
data("florentine_m")
#Delete Pucci family (isolated)
florentine_m <- delete_vertices(florentine_m,which(degree(florentine_m)==0))
```

```{r plot,fig.height=5,fig.width=5,fig.align='center'}
#plot the graph
set.seed(111)
plot(florentine_m,
     vertex.label.cex=V(florentine_m)$wealth*0.01,
     vertex.label.color="black",
     vertex.color="white",
     vertex.frame.color="gray")
```

________________________________________________________________________________

## Most central family

```{r cent}
cent.df <- data.frame(
  degree=degree(florentine_m),
  betweenness=betweenness(florentine_m),
  closeness=closeness(florentine_m),
  eigenvector=eigen_centrality(florentine_m)$vector,
  subgraph=subgraph_centrality(florentine_m))

# most central family according to the 5 indices
V(florentine_m)$name[apply(cent.df,2,which.max)]
```

In all cases, the Medici are considered to be the most central family. However,
we can also find an index that ranks other families on top.

```{r cent_new}
#odd subgraph centrality
sc_odd <- florentine_m %>% 
  indirect_relations(type = "walks",FUN = walks_exp_odd) %>% 
  aggregate_positions(type = "self")

#family with highest score 
V(florentine_m)$name[which.max(sc_odd)]
```

```{r ni}
P <- neighborhood_inclusion(florentine_m)
comparable_pairs(P)
```

```{r dom_graph,fig.height=5,fig.width=5,fig.align='center'}
d <- dominance_graph(P)
V(d)$name <- V(florentine_m)$name
set.seed(113)
plot(d,vertex.label.color="black",
     vertex.color="white",
     vertex.frame.color="gray",
     edge.arrow.size=0.5)
```

```{r probs}
res <- exact_rank_prob(P,names=V(florentine_m)$name)
str(res)
```
There are `r format(res$lin.ext,scientific = F,big.mark = ",")` different possibilities to rank
the families.

```{r likely_most_central}
top_rank_prob <- res$rank.prob[,15]
names(top_rank_prob) <- V(florentine_m)$name
round(top_rank_prob[top_rank_prob>0],3)
```
Highest prob for Strozzi

________________________________________________________________________________

## Hypothesis

Can proximity explain wealth?

Index approach
```{r dist_vs_wealth}

#Closeness
c_C <- florentine_m %>% 
  indirect_relations(type="geodesic") %>% 
  aggregate_positions(type="invsum")

#harmonic closeness
c_HC <- florentine_m %>% 
  indirect_relations(type="geodesic",FUN=dist_inv) %>% 
  aggregate_positions(type="sum")

cor(c_C,V(florentine_m)$wealth,method="kendall")
cor(c_HC,V(florentine_m)$wealth,method="kendall")
```

There doesn't seem to be much evidence, yet there are several additional distance based
indices!

```{r dist_vs_wealth2}
#residual closeness (Dangalchev,2006)
c_RC <- florentine_m %>% 
  indirect_relations(type="geodesic",FUN=dist_2pow) %>% 
  aggregate_positions(type="sum")

#integration centrality (Valente & Foreman, 1998)
dist_integration <- function(x){
  x <- 1 - (x - 1)/max(x)
}
c_IN <- florentine_m %>% 
  indirect_relations(type="geodesic",FUN=dist_integration) %>% 
  aggregate_positions(type="sum")

cor(c_RC,V(florentine_m)$wealth,method="kendall")
cor(c_IN,V(florentine_m)$wealth,method="kendall")
```

There are also some indices with parameters to maximize correlation.

```{r distalpha_wealth,warning=FALSE}
#generalized closeness (Agneessens et al.,2017) (alpha>0)
alpha <- c(seq(0.01,0.99,0.01),seq(1,10,0.1))
scores <- 
sapply(alpha,function(x)
    florentine_m %>% 
      indirect_relations(type="geodesic",FUN=dist_dpow,alpha=x) %>% 
      aggregate_positions(type="sum")
)
cors <- apply(scores,2,
              function(x)cor(x,V(florentine_m)$wealth,method="kendall"))

res_gc <- c(max(cors),alpha[which.max(cors)])

#decay centrality (Jackson, 2010) (alpha in [0,1])
alpha <- seq(0.01,0.99,0.01)
scores <- 
sapply(alpha,function(x)
  florentine_m %>% 
    indirect_relations(type="geodesic",FUN=dist_powd,alpha=x) %>% 
    aggregate_positions(type="sum")
)
cors <- apply(scores,2,
              function(x)cor(x,V(florentine_m)$wealth,method="kendall"))

res_dc <- c(max(cors),alpha[which.max(cors)])
```

The highest correlation for generalized closeness is `r res_gc[1]` achieved for 
alpha= `r res_gc[2]`  

The highest correlation for decay centrality is `r res_dc[1]` achieved for 
alpha=`r res_dc[2]`